---
title: "Clustering of Spotify data to Infer Relationships of Genres and Show US Trends Over Time"
author: "Angel Sarmiento"
subtitle: Florida Polytechnic University
output:
  pdf_document: default
  html_notebook: default
url_color: blue
---

```{r, include = FALSE}
library(tinytex)
```


## Summary

Spotify is a music streaming service that provides millions of people with virtually unlimited amounts of music spanning across multiple genres and time periods. All a person has to do to access this expansive catalogue is to sign up and start listening. Lots of data and metadata characterize music and a variety of very useful statistical information can be derived from such a large database of information. The objective of this project is to leverage this data to answer some questions about the patterns of music trends and to generate a model that can sufficiently infer the relationships of genres, artists, and albums.  
\
Another function of this work would be to create a spectrum of "feelings" for some individual songs that will show the emotional qualities of them and then to assign them to specific regions where those songs are available. A different, more mechanical way of showing this would be to analyze how the composition of music has changed over time by utilizing the *key centers* of songs, tempo, number of keys, and *instrumentalness*. *Keys* being the specific group of notes that establish the tonality of a song and *instrumentalness* being how instrument heavy the composition of a song is.  


The existing questions posed can be listed as:
\begin{itemize}
\item What relationships are there between different genres/styles of music?  
\item What kinds of emotions are displayed in music and how does that change over time?  
\item Which regions in the world listen to which genres and emotional qualities of music, and can that be related with seasonality at all?  
\item How has music composition change over time? 
\end{itemize}

## Resources

The primary resource for this project is the ['spotifyr'] package, an wrapper for the Spotify web API which can be found with the following R commands
```{r, eval = FALSE}
install.packages("spotifyr")
library("spotifyr")
```
Other resources include 
\begin{itemize}
\item \href{http://paulelvers.com/post/emotionsineuropeanmusic/}{{\color{blue}{\underline{Sentiment analysis of musical taste: a cross-European comparison
}}}}
\item A paper titled \href{https://www.math.vu.nl/~sbhulai/papers/paper-vandenhoven.pdf}{{\color{blue}{\underline{Analyzing Spotify Data}}}} written by Jeroen van den Hoven from VU University.
\end{itemize}
